%& -shell-escape
\documentclass{llncs}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[british]{babel}
\usepackage{fixltx2e}
\usepackage{xparse}
\usepackage{booktabs}
\usepackage{pythontex}
\usepackage[defblank]{paralist}

\usepackage[fulladjust]{marginnote}

\usepackage[natbib,style=lncs]{biblatex}
\addbibresource{otrmsg.bib}
\addbibresource{surveillance.bib}
\addbibresource{crypto.bib}
\addbibresource{otpkx.bib}
\addbibresource{nfc.bib}

\usepackage{csquotes}
\MakeBlockQuote{<}{|}{>}
\EnableQuotes{}

\usepackage{filecontents}

\usepackage[algebra,crypto,surveillance,std,stdmaths]{libbib}

\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{amssymb}
\DeclareMathOperator{\com}{\textsc{com}}

\usepackage[binary,amssymb]{SIunits}

\usepackage[plain]{algorithm}
\usepackage[noend]{algpseudocode}
\NewDocumentCommand{\alg}{m}{%
  \textsc{#1}%
}

\usepackage[capitalize]{cleveref}
\crefname{chapter}{Chap.}{Chap.}
\Crefname{chapter}{Chapter}{Chapters}
\crefname{section}{Sect.}{Sect.}
\Crefname{section}{Section}{Sections}
\crefname{figure}{Fig.}{Fig.}
\Crefname{figure}{Figure}{Figures}
\crefname{table}{Tab.}{Tab.}
\Crefname{table}{Table}{Tables}
\crefname{definition}{Def.}{Def.}
\Crefname{definition}{Definition}{Definitions}
\crefname{theorem}{Thm.}{Thm.}
\Crefname{theorem}{Theorem}{Theorems}
\crefname{lemma}{Lem.}{Lem.}
\Crefname{lemma}{Lemma}{Lemmas}
\crefname{corollary}{Cor.}{Cor.}
\Crefname{corollary}{Corollary}{Corollaries}
\crefname{proposition}{Prop.}{Prop.}
\Crefname{proposition}{Proposition}{Propositions}

\usepackage{msc}
\setmscvalues{small}
\setlength{\instdist}{3cm}
%\setlength{\instwidth}{1.6cm}
%\setlength{\leftnamedist}{0cm}
\setlength{\topheaddist}{\bottomfootdist}
%\setlength{\bottomfootdist}{0cm}
\setlength{\firstlevelheight}{20pt}
\setlength{\levelheight}{20pt}

\pagestyle{plain}

\spnewtheorem{protocol}{Protocol}{\bfseries}{\normalfont}

\includeonly{otpkx-content}
\begin{document}
\title{%
  Towards Perfectly Secure and Deniable Communication Using an NFC-Based 
  Key-Exchange Scheme
}
\author{%
  Daniel Bosk\inst{1}
  \and
  Martin Kjellqvist\inst{2}
  \and
  Sonja Buchegger\inst{1}
}
\institute{%
  School of Computer Science and Communication,\\
  KTH Royal Institute of Technology, SE-100\,44 Stockholm\\
  Email: \email{\{dbosk,buc\}@kth.se}
  \and
  Department of Information and Communication Systems,\\
  Mid Sweden University, SE-851\,70 Sundsvall\\
  Email: \email{martin.kjellqvist@miun.se}
}

\maketitle
\begin{abstract}
  In this paper we first analyse the possibility for deniability under a strong 
  adversary, who has an Internet-wide transcript of the communication.
  Secondly, we present a scheme which provides the desirable properties of 
  previous messaging schemes, but with stronger deniability under the new 
  adversary model.
  Our scheme requires physical meetings for exchanges of large amounts of 
  random key-material via near-field communication and later uses this random 
  data to key a one-time pad for text-messaging.
  We prove the correctness of the protocol and, finally, we evaluate the 
  practical feasibility of the suggested scheme.

  \keywords{%
    Deniability,
    Deniable Encryption,
    Authenticated Encryption,
    Perfect Secrecy,
    Off-the-Record,
    Key-Exchange,
    Near-Field Communication,
    Surveillance.
  }
\end{abstract}

\acresetall

\include{otpkx-content}

\printbibliography
\end{document}
